---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nettskjemar

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/LCBC-UiO/nettskjemar.svg?branch=master)](https://travis-ci.org/LCBC-UiO/nettskjemar)
[![Codecov test coverage](https://codecov.io/gh/LCBC-UiO/nettskjemar/branch/master/graph/badge.svg)](https://codecov.io/gh/LCBC-UiO/nettskjemar?branch=master)
<!-- badges: end -->

The goal of nettskjemar is to ...


Install the development version from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("LCBC-UiO/nettskjemar")
```

## Setting up the authentication
The is a [wiki](https://github.com/LCBC-UiO/nettskjemar/wiki/Setting-up-authentication) on setting up the authentication to access the nettskjema through the api.
This will need to be completed before you can access any other features of this package.

## Example

Currently there are very few functions actually working in the package.
The most important, is the function that will download nettskjema submission data and return them as a tibble (data.frame). 
This needs only the _id_ of a nettskjema, which can be found in the last part of the nettskjema url.

```{r example, eval=FALSE}
library(nettskjema)

nettskjema_get_data(nettskjema_id)
```

If you do not have the codebook activated, or you want to use the full answers to questions rather than the coded ones, you can toggle off code book answers:

```{r, eval=FALSE}
nettskjema_get_data(nettskjema_id, use_codebook = FALSE)
```


## Documentation
Package documentation can be found on the associated [GitHub pages](https://lcbc-uio.github.io/nettskjemar/), among other documentation on [how to create the API acces user and token](https://lcbc-uio.github.io/nettskjemar/articles/auth_setup.html).

