---
title: "Attachments: Download and Save"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Attachments: Download and Save}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| label: atch-1
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = FALSE,
  fig.path = "static"
)

vcr::setup_knitr(prefix = "atch-")
```

# Overview

Nettskjema allows users to upload attachments with their form submissions. This vignette explains how to use the functions provided in this package to list, retrieve, and save these attachments efficiently.

# Listing Attachments

Before downloading any attachments, it's useful to list the available files.

## Listing attachments for a specific submission

Each submission ID may have associated attachments. Use `ns_list_submission_attachments()` to retrieve information about the available attachments.

```{r}
#| label: atch
#| cassette: sub

library(nettskjemar)
# Example: List attachments for a specific submission
submission_id <- 27685292
ns_list_submission_attachments(submission_id)
```

This function returns a `data.frame` with details such as the attachment ID, original filename, and standardized filename.

## Listing all attachments for a form

To list all attachments across multiple submissions for a form, use `ns_list_form_attachments()`.

```{r}
#| label: atch-all
#| cassette: all

# Example: List attachments for an entire form
formid <- 123823
ns_list_form_attachments(formid)
```

# Downloading Attachments

Once you have identified the attachments, you can download them using `ns_get_attachment()`.

```r
#| label: atch-dl1

# Example: Download a single attachment
attachment_id <- "12345"
ns_get_attachment(attachment_id, path = "downloaded_file.pdf")
```

If `path` is not provided, the function returns the response object without saving the file.

# Saving All Attachments

## For a Single Submission

Use `ns_get_submission_attachments()` to download and save all attachments associated with a single submission.

```r
#| label: atch-dl2

# Example: Save all attachments for a submission
ns_get_submission_attachments(submission_id)
```

This function supports two filename conventions:

- **"standardized"** (default): Creates filenames based on submission ID and a counter.
- **"original"**: Uses the filenames as they were uploaded.

To save attachments using the original filenames:

```r
#| label: atch-dl3

ns_get_submission_attachments(submission_id, filenames = "original")
```

You can also specify an output directory:

```r
#| label: atch-dl4

ns_get_submission_attachments(submission_id, output_dir = "my_attachments/")
```

## For an Entire Form

To download and save all attachments from a form:

```r
#| label: atch-dl5

ns_get_form_attachments(formid)
```

As with submission attachments, you can specify a filename format and output directory:

```r
#| label: atch-dl6

ns_get_form_attachments(
  formid,
  filenames = "original",
  output_dir = "all_attachments/"
)
```

# Summary

This vignette demonstrated how to:
- List attachments for submissions and forms.
- Retrieve individual attachments.
- Save attachments using different naming conventions.

By using these functions, you can efficiently manage Nettskjema file attachments within your R workflow.

