---
title: "Using nettskjemar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using nettskjemar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Nettskjemar connects to version 2 of the [nettskjema api](https://nettskjema.no/), and the main functionality here is to download data from a form into R.
Once you have [created a nettskjema api user](auth_setup.html), and set up your Renvironment locally, you can start accessing your forms.

## Download submissions
Perhaps at the core of nettskjemar is the ability to download submission answers to a form into a tibble (variation of a data.frame). 

```{r}
library(nettskjemar)

nettskjema_get_data(110000)
```

There are many arguments that can be set that give you control over the data extraction. 
By default, the data as set by the codebook is retrieved by `use_codebook = TRUE`, this can be set to `FALSE` which would retrieve the full-text information.
If the codebook has not been set up, initial download will fail because there is none, and the user must toggle this themselves in the nettskjema-portal. 

```{r}
nettskjema_get_data(110000, use_codebook = FALSE)
```



